---
title: "Site Traffic Analytics"
toc: true
toc-location: right
code-fold: true
code-summary: "Show code"
execute:
  freeze: true
  echo: true
  warning: false
  message: false
---

This page analyzes the real traffic to this portfolio site. When I built the site, I added a Google Analytics 4 tag, linked the property to BigQuery, and turned on the daily export. Every visit now flows from GA4 into a BigQuery table that I query directly from R. This page is that pipeline working end to end: GA4 collects the events, BigQuery stores them, and R and Quarto turn them into the analysis below.

::: {.callout-note appearance="simple"}
**A note on the data.** This site is newly launched, so the numbers here reflect early traffic rather than months of history. I built the page on my own live data on purpose. The point is to show the full GA4 to BigQuery to R workflow on real events I collected myself, and the analysis will only get richer as the export accumulates more days.
:::

```{r}
#| label: setup
library(bigrquery)
library(DBI)
library(tidyverse)
library(gt)


project_id  <- "ibm-6530-487205"
ga4_dataset <- "analytics_543432183"

events_tbl <- paste0("`", project_id, ".", ga4_dataset, ".events_*`")

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_id,
  billing = project_id
)

# A small house style so every chart on the page looks consistent
bar_fill <- "#2c6e91"
theme_set(theme_minimal(base_size = 12))
```

## Headline metrics

The quickest read on the site: how many people have visited, how many sessions they started, and how many pages they looked at over the period I have data for.

```{r}
#| label: headline
sql_headline <- paste0("
  SELECT
    MIN(PARSE_DATE('%Y%m%d', event_date)) AS first_day,
    MAX(PARSE_DATE('%Y%m%d', event_date)) AS last_day,
    COUNT(DISTINCT user_pseudo_id) AS total_users,
    COUNT(DISTINCT CONCAT(
      user_pseudo_id,
      CAST((SELECT value.int_value FROM UNNEST(event_params)
            WHERE key = 'ga_session_id') AS STRING)
    )) AS total_sessions,
    COUNTIF(event_name = 'page_view') AS total_page_views
  FROM ", events_tbl, "
")

headline <- dbGetQuery(con, sql_headline)

headline |>
  transmute(
    `Date range`   = paste(format(first_day, "%b %d"), "to", format(last_day, "%b %d, %Y")),
    `Total users`  = total_users,
    `Sessions`     = total_sessions,
    `Page views`   = total_page_views
  ) |>
  gt() |>
  tab_header(title = "Portfolio traffic at a glance") |>
  fmt_number(columns = c(`Total users`, `Sessions`, `Page views`), decimals = 0) |>
  cols_align(align = "center") |>
  tab_options(table.font.size = px(15), heading.title.font.size = px(18))
```

## Users and sessions over time

Tracking users and sessions by day shows whether traffic is steady or spiky. With a young site, this is mostly a baseline to build on as more days of export come in.

```{r}
#| label: over-time
#| output: asis
sql_daily <- paste0("
  SELECT
    PARSE_DATE('%Y%m%d', event_date) AS date,
    COUNT(DISTINCT user_pseudo_id) AS users,
    COUNT(DISTINCT CONCAT(
      user_pseudo_id,
      CAST((SELECT value.int_value FROM UNNEST(event_params)
            WHERE key = 'ga_session_id') AS STRING)
    )) AS sessions
  FROM ", events_tbl, "
  GROUP BY date
  ORDER BY date
")

daily <- dbGetQuery(con, sql_daily)

# With only one day of export, a trend line is not meaningful yet, so show a
# short note instead. Once a few daily tables have accumulated, the real chart
# renders automatically.
if (dplyr::n_distinct(daily$date) >= 2) {
  p <- daily |>
    pivot_longer(c(users, sessions), names_to = "metric", values_to = "count") |>
    ggplot(aes(date, count, color = metric)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    scale_color_manual(values = c(users = "#2c6e91", sessions = "#e76f51"),
                       labels = c("Users", "Sessions")) +
    scale_x_date(date_labels = "%b %d") +
    scale_y_continuous(limits = c(0, NA), breaks = scales::breaks_pretty()) +
    labs(x = NULL, y = NULL, color = NULL,
         title = "Daily users and sessions") +
    theme(legend.position = "top")
  print(p)
} else {
  cat('::: {.callout-note appearance="simple"}\n',
      '**This chart builds as data accumulates.** With a single day of export so far, ',
      'a day-over-day trend is not yet meaningful. It will fill in on its own as more ',
      'daily GA4 exports land.\n',
      ':::\n', sep = '')
}
```

## Where my visitors come from

This groups traffic by source and medium, which is the GA4 way of describing how someone arrived (a direct visit, a link from another site, a search, and so on). For a portfolio I share by link, I expect a lot of direct and referral traffic, and this confirms it.

```{r}
#| label: sources
sql_sources <- paste0("
  SELECT
    COALESCE(traffic_source.source, '(direct)') AS source,
    COALESCE(traffic_source.medium, '(none)')   AS medium,
    COUNT(DISTINCT user_pseudo_id) AS users
  FROM ", events_tbl, "
  GROUP BY source, medium
  ORDER BY users DESC
")

sources <- dbGetQuery(con, sql_sources)

sources |>
  gt() |>
  tab_header(title = "Traffic by source and medium") |>
  cols_label(source = "Source", medium = "Medium", users = "Users") |>
  fmt_number(columns = users, decimals = 0) |>
  tab_options(table.font.size = px(14))
```

```{r}
#| label: sources-plot
#| output: asis
# A bar chart with a single channel just shows one full-width bar, so when
# there is only one source/medium, show a short note instead.
if (nrow(sources) >= 2) {
  p <- sources |>
    mutate(channel = paste(source, medium, sep = " / ")) |>
    slice_max(users, n = 8) |>
    ggplot(aes(reorder(channel, users), users)) +
    geom_col(fill = bar_fill, width = 0.7) +
    coord_flip() +
    scale_y_continuous(breaks = scales::breaks_pretty()) +
    labs(x = NULL, y = "Users", title = "Top acquisition channels")
  print(p)
} else {
  cat('::: {.callout-note appearance="simple"}\n',
      '**One channel so far.** Every visit to date arrived directly through a shared link, ',
      'so there is a single acquisition channel to report. This view becomes more useful ',
      'once referral, search, or social traffic starts to appear.\n',
      ':::\n', sep = '')
}
```

## Top pages

Which pages people actually open tells me what the site is doing well and where attention goes first. I cleaned the full URLs down to just the path so the table is easy to read.

```{r}
#| label: top-pages
sql_pages <- paste0("
  SELECT
    (SELECT value.string_value FROM UNNEST(event_params)
       WHERE key = 'page_location') AS page_location,
    COUNT(*) AS views
  FROM ", events_tbl, "
  WHERE event_name = 'page_view'
  GROUP BY page_location
  ORDER BY views DESC
  LIMIT 15
")

pages <- dbGetQuery(con, sql_pages)

pages |>
  mutate(
    path = sub("^https?://[^/]+", "", page_location),
    path = ifelse(path == "" | is.na(path), "/", path)
  ) |>
  select(path, views) |>
  gt() |>
  tab_header(title = "Most viewed pages") |>
  cols_label(path = "Page", views = "Views") |>
  fmt_number(columns = views, decimals = 0) |>
  tab_options(table.font.size = px(14))
```

## Devices and geography

A quick look at the kind of device people browse on and where they are. For a portfolio that hiring managers might open on a phone between meetings, the device split is genuinely useful to know.

```{r}
#| label: devices
sql_devices <- paste0("
  SELECT
    device.category AS device_category,
    COUNT(DISTINCT user_pseudo_id) AS users
  FROM ", events_tbl, "
  GROUP BY device_category
  ORDER BY users DESC
")

devices <- dbGetQuery(con, sql_devices)

devices |>
  filter(!is.na(device_category)) |>
  ggplot(aes(reorder(device_category, users), users)) +
  geom_col(fill = bar_fill) +
  coord_flip() +
  labs(x = NULL, y = "Users", title = "Users by device category")
```

```{r}
#| label: geography
sql_geo <- paste0("
  SELECT
    COALESCE(geo.country, '(unknown)') AS country,
    COUNT(DISTINCT user_pseudo_id) AS users
  FROM ", events_tbl, "
  GROUP BY country
  ORDER BY users DESC
  LIMIT 10
")

geo <- dbGetQuery(con, sql_geo)

geo |>
  gt() |>
  tab_header(title = "Visitors by country") |>
  cols_label(country = "Country", users = "Users") |>
  fmt_number(columns = users, decimals = 0) |>
  tab_options(table.font.size = px(14))
```

## How this page was built

The workflow behind everything above is the same pipeline marketing teams use to move analytics out of the GA4 interface and into a reproducible report:

1. **Collect.** A GA4 tag on every page of this site records events as visitors browse.
2. **Store.** GA4 exports those events to BigQuery once a day, where each day becomes its own `events_` table.
3. **Query.** I pull the data into R with the `bigrquery` package and write the analysis in BigQuery SQL.
4. **Report.** R, `gt`, and `ggplot2` turn the query results into the tables and charts on this page, all rendered through Quarto.

The natural complement to this is a Looker Studio dashboard connected to the same GA4 property, which gives a live, always-current view for quick checks. This page is the reproducible, code-driven side of that same data.

```{r}
#| label: cleanup
dbDisconnect(con)
```
 